The KMOS Deep Survey (KDS): dynamical measurements of $z\sim3.5$ main sequence galaxies\thanks{Based on observations obtained at the Very Large Telescope of the European Southern Observatory. Programme IDs: }

### 2017-03-10 paper.tex
draft version 2 sent to co-authors

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
%\author[O.J. Turner et al.]{
%O. J. Turner,$^{1,2,\thanks{E-mail: turner@roe.ac.uk (OJT)}}$
%M. Cirasuolo,$^{1,2}$
%C. M. Harrison,$^{2,3}$
%J. Dunlop,$^{1}$
%R. J. McLure,$^{1}$\newauthor
%A. M. Swinbank$^{4,3}$
%H. L. Johnson$^{3,4}$
%\\
%$^{1}$SUPA\thanks{Scottish Universities Physics Alliance}, Institute for Astronomy, University of Edinburgh, Royal Obervatory, Edinburgh EH9 3HJ\\
%$^{2}$European Southern Observatory, Karl-Schwarzschild-Str. 2, 85748 Garching b. M{\"u}nchen, Germany\\
%$^{3}$Centre for Extragalactic Astronomy, Durham University, South Road, Durham, DH1 3LE, U.K.\\
%$^{4}$Institute for Computational Cosmology, Durham University, South Road, Durham, DH1 3LE, U.K.
%}

@@

% Abstract of the paper
\begin{abstract}
We present dynamical measurements for 77 main sequence star-forming galaxies at $z\sim3.5$ in the mass range $9.5 < log(M_{\star}[M_{\odot}]) < 10.5$ from the KMOS (K-band Multi-Object Spectrograph) Deep Survey (KDS).
These measurements constitute the largest sample at $z > 3$ utilising Integral-Field Spectroscopy (IFS), and provide robust constraints on the internal dynamics, via the intrinsic velocity dispersion $\sigma_{int}$ and rotation velocities, $V_{C}$, of galaxies in the high-redshift universe.
We have made careful comparisons to other IFS samples which have been selected consistently, span a wide redshift baseline and use ionised gas emission to trace galaxy dynamics. 
The KDS mean dispersion value is $\sigma_{int} = 70.8^{+3.3}_{-3.1} kms^{-1}$, in agreement with the increase of $\sigma_{int}$ with redshift reported across the comparison samples.
However we find that only $34 \pm 8\%$ of our galaxies are rotationally-dominated ($V_{C}/\sigma_{int}$ > 1), with the sample average $V_{C}/\sigma_{int}$ value much less than at lower redshift.
The rotation dominated KDS galaxies show no clear offset from the local rotation velocity-stellar mass (i.e. $V_{C}-M_{\star}$) relation (the inverse of the stellar mass Tully-Fisher relation), although a smaller fraction of the galaxies sit on the relation due to the increase in the dispersion dominated fraction.
The galaxies show a decrease in specific angular momentum (angular momentum divided by stellar mass) of $\sim0.6-0.7$dex in comparison to $z=0$ disk galaxies, with this decline the result of the gradual size evolution of the star-forming population over cosmic time.
Due to these more compact sizes, the KDS galaxies show a paucity of virial mass in comparison to stellar mass, which may be explained through an additional virial component originating in random motions, as traced by $\sigma_{int}$.
These observations are in line with a simple equilibrium model picture, in which random motions are boosted in high-redshift galaxies by the increase in gas fractions, accretion efficiency, specific star-formation rate and stellar feedback.
\end{abstract}

@@

In the hierarchical collapse scenario \citep{Rees1977,Fall1980}, matter density perturbations, which originate from some set of cosmological initial conditions, grow over time before condensing out of the general expansion to form dark matter halos \citep{Silk1968,Peebles1970}.
It is at the centre of these halos that baryons collect, cool and begin to form stars (\citealt{Birnboim2003,Keres2005,Dekel2006} and see e.g. \citealt{Mo2010} for a recent review), with these systems constituting the first galaxies.
One crucial ingredient shaping the subsequent evolution of this system is the transfer of angular momentum from the collapsing gas, as a protogalaxy which is rich in angular momentum is likely to form a turbulent, rotating disk, whereas if the angular momentum is low the formation of a spherical system dominated by random motions will ensue \citep[e.g.][]{Fall1983}.

It has been observationally determined that the galaxy population is bimodal in many physical properties at each redshift slice at which it is studied \citep[e.g. as described in][]{Dekel2006}, with a preference for the most massive galaxies to lie in the red sequence, characterised by red optical colours, low Star Formation Rates (SFRs) and spherical morphologies, and less massive galaxies in the blue sequence, characterised by blue colours, high star formation rates and disky morphologies. 
For these blue, star-forming galaxies (SFGs) there is a roughly linear correlation between SFR and stellar mass ($M_{\star}$) \citep[e.g.][]{Daddi2007,Noeske2007,Elbaz2007}, in the sense that galaxies which have already accumulated a larger stellar population tend to have higher SFRs.
This correlation, or `main sequence' underpins what is known as the equilibrium model, in which the SFR of galaxies is regulated by the availability of gas, with outflows and accretion events sustaining the galaxy gas reservoirs in a rough equilibrium as the galaxy evolves \citep[e.g.][]{Dave2012,Lilly2013,Saintonge2013}.
The main sequence has been studied comprehensively, using multi-wavelength SFR tracers, between $0 < z < 3$ \citep[e.g.][]{Rodighiero2011,Karim2011,Whitaker2012,Behroozi2013b,Whitaker2014,Rodighiero2014,Speagle2014,Pannella2014,Sobral2014,Sparre2015,Lee2015,Schreiber2015,Renzini2015,Nelson2016}, finding evolution of the relation towards higher SFRs at fixed $M_{\star}$ with increasing redshift, reflecting the increase of the cosmic Star Formation Rate Density (SFRD) in this redshift range \citep{Madau_2014}.
At each redshift slice, galaxies on the main sequence are undergoing secular evolution regulated by their gas reservoirs and so selecting such populations offers the chance to explore the evolution of the physical properties of typical SFGs across cosmic time.
This assumes that high-redshift main sequence galaxies are the progenitors of their lower redshift counterparts, which may not be the case \citep[e.g.][]{Gladders2013,Kelson2014,Abramson2016b} and also assumes that we can learn about galaxy evolution (i.e. how individual galaxies develop in physical properties over time) by studying the mean properties of populations at different epochs.
The picture is complicated by the addition of major and minor galaxy mergers which can rapidly change the physical properties of galaxies \citep[e.g.][]{Toomre1977,Lotz2008,Conselice2011,Conselice2014} and the relative importance of in-situ, secular stellar mass growth vs. stellar mass aggregation via mergers is the subject of much work involving both observations and simulations \citep[e.g.][]{Robaina2009,Kaviraj2012,Stott2013,Lofthouse2016,Qu2016}. 
To account for the growing number density of quiescent galaxies from z$\sim 2.5$ to the present day \citep[e.g.][]{Bell2004,Faber2007,Brown2007,Ilbert2010,Brammer2011,Muzzin2013,Buitrago2013} there must also be some process which shuts off star-formation within main sequence galaxies (i.e. quenching) which, to explain observations, is a function of both mass and environment \citep{Peng2010,Darvish2016}.

Recent cosmological volume simulations are able to resolve the complex interplay of baryonic processes which are at work as galaxies evolve, and can track the development of individual galaxies from early stages, to maturity and through quenching \citep{Dubois2014a,Vogelsberger2014b,Schaye2015}.
Observations can aid the predictive power of such simulations by providing constraints on the evolving physical properties of galaxy populations which must be reproduced across cosmic time.
The observed dynamical properties of galaxies contain information about the transfer of angular momentum between dark matter halos and the baryons and the dissipation of this angular momentum thereafter (through gravitational collpase, mergers and outflows \citealt{Fall1983,Romanowsky2012,Fall2013}), constituting an important set of quantities for simulators to attempt to match.
Developments in both Integral-Field Spectroscopy (IFS) instrumentation and data analysis tools over the last decade have led to the observation of two-dimensional velocity and velocity dispersion fields for large samples of galaxies of different morphological types, spanning a wide redshift range \citep[e.g.][]{Sarzi2005,Flores2006,Epinat2008,ForsterSchreiber2009,Cappellari2011,Gnerucci2011,Epinat2012,Croom2012,Swinbank2012,Swinbank2012a,Bundy2015,Wisnioski2015,Stott2016,Harrison2017,Swinbank2017} and when interpreted in tandem with high-resolution imaging data from the Hubble Space Telescope ({\em {\em HST}}) these data provide exquisite information about the range of physical processes which are driving galaxy evolution.
In particular, in recent years the multiplexing capabilities of KMOS have allowed for IFS kinematic observations for large galaxy samples to build up over much shorter timescales \citep{Sobral2013,Wisnioski2015,Stott2016,Mason2016,Harrison2017} providing an order of magnitude boost in statistical power over previous high-redshift campaigns.

SFGs generally appear to be increasing in random motions with increasing redshift between $0 < z < 3$, as traced by their observed velocity dispersions, $\sigma_{obs}$ \citep{Genzel2008,ForsterSchreiber2009,Law2009,Cresci2009,Gnerucci2011,Epinat2012,Kassin2012,Green2014,Wisnioski2015,Stott2016}.
This has been explained in terms of increased `activity' in galaxies during and before the global peak in cosmic SFRD \citep{Madau_2014}, in the form of higher specific star formation rates (sSFR) \citep{Wisnioski2015}, larger gas reservoirs \citep{Law2009,ForsterSchreiber2009,Wisnioski2015,Stott2016}, more efficient accretion \citep{Law2009}, increased stellar feedback from supernovae \citep{Kassin2012} and turbulent disk instabilities \citep{Law2009,Bournaud2007,Bournaud2016}, all of which combine to increase $\sigma_{obs}$ and complicate its interpretation.
There is also an increasing body of work measuring the relationship between the observed maximum rotation velocity of a galaxy, a tracer for the total dynamical mass, and its stellar mass, known as the stellar mass Tully-Fisher Relation (smTFR) \citep{Tully1977}, with surveys reporting disparate results for the evolution of this relation with redshift \citep[e.g.][]{Puech2008,Miller2011,Gnerucci2011,Swinbank2012,Tiley2016a,Harrison2017}.
In both cases, systematic differences in measurement and modelling techniques at high-redshift, especially with regards to beam-smearing corrections, combine with our poor understanding of progenitors and descendants to blur the evolutionary picture which these surveys paint.

@@

\subsubsection{GOODS-S}\label{subsubsec:sample_selection_goods}
Two of the FIELD environment pointings (GOODS-S-P1;GOODS-S-P2; see table \ref{tab:pointings}) are selected within GOODS-S region \citep{Guo2013}; accessible from the VLT and with excellent multi-wavelength coverage, including deep {\em HST WFC3 F160W} imaging with $ 0.06^{\prime\prime}$ pixel scale and $ \sim 0.2^{\prime\prime}$ PSF, which is well suited for constraining galaxy morphology \citep{Grogin2011,Koekemoer2011}.
We select targets from the various spectroscopic campaigns which have targetted GOODS-S, including measurements from VIMOS \citep{Balestra2010,Cassata2014}, FORS2 \citep{Vanzella2005,Vanzella2006,Vanzella2008} and both LRIS and FORS2 as outlined in \cite{Wuyts2009}.
These targets must be within the redshift range $3 < z < 3.8$, have high spectral quality (as quantified by the VIMOS redshift flag equal `3' or `4', and the FORS2 quality flag equal `A') and we carefully exclude those targets for which the [O~{\sc III}]$\lambda$5007 or H$\beta$ emission lines, observable in the K-band at these redshifts, would be shifted into a spectral region plagued by strong OH emission.
The galaxies which remain after imposing these criteria are distributed across the GOODS-S field, and we select two regions where $\sim 20$ targets can be allocated to the KMOS IFUs (noting that the IFUs can patrol a $7.2^{\prime}$ diameter patch of sky during a single pointing).
In GOODS-S-P1 and GOODS-S-P2 we observe 20 and 17 galaxies respectively.\footnote{We note that the number of observed galaxies in GOODS-S-P2 does not include two targets which were later found to have $z < 0.5$.}

\subsubsection{SSA22}\label{subsubsec:sample_selection_ssa}
A single CLUSTER environment pointing (SSA22-P1; see table \ref{tab:pointings}) is selected from the SSA22 field, \citep{Steidel1998,Steidel2000,Steidel2003,Shapley2003}, which, as mentioned above, is an overdensity of LBG candidates at $z \sim 3.09$.
Thousands of spectroscopic redshifts have been confirmed for these LBGs with follow up observations using LRIS \citep{Shapley2003,Nestor2013}.
A combination of deep {\em B,V,R} band imaging with the Subaru Suprime-Cam \citep{Matsuda2004}, deep narrow band imaging at 3640$\AA$ \citep{Matsuda2004} and at 4977$\AA$ \citep{Nestor2011} and archival {\em HST ACS} and {\it WFC3} provides ancillary data in excellent support of integral field spectroscopy, albeit over a shorter wavelength baseline and with shallower exposures than in the GOODS-S field.
Fortunately at $z \sim 3.09$ the [O~{\sc III}]$\lambda$5007 line is shifted into a region of the K-band which is free from OH features and so for the CLUSTER environment pointing we fill the KMOS IFUs with galaxies located towards the centre of the SSA22 protocluster (SSA22-P1).

We also add a further FIELD environment pointing to the south of the main SSA22 spatial overdensity where the density of galaxies is typical (SSA22-P2).
In SSA22-P1 and SSA22-P2 we observe 19 and 21 galaxies respectively.
In summary we have chosen three FIELD environment pointings and a single CLUSTER environment pointing across GOODS-S and SSA22, comprising a total of 77 galaxies, as described in table \ref{tab:pointings}. \footnote{Additional pointings in the COSMOS and UDS fields were originally scheduled as part of the GTO project, however 50$\%$ of the observing time was lost to bad weather during these visitor mode observations.}

\subsection{Observations and data reduction}\label{subsubsec:observations_and_dr}
Our data for the 77 KDS targets were observed using KMOS.
KMOS is a second generation IFS mounted at Nasmyth focal plane on UT1 at the VLT.
The instrument has 24 moveable pickoff arms, each with an integrated IFU, which patrol a region 7.2$^{\prime}$ in diameter on the sky, providing considerable flexibility when selecting sources for a single pointing.
The light from each set of 8 IFUs is dispersed by a single spectrograph and recorded on a 2k$\times$2k Hawaii-2RGHgCdTe near-IR detector, so that the instrument is comprised of three effectively independent modules.
Each IFU has 14$\times$14 spatial pixels which are 0.2$^{\prime\prime}$ in size, and the central wavelength of the K-band filter has spectral resolution of $R \sim 4200$.
\subsubsection{Observations}\label{subsubsec:Obs}
To achieve the science goals of the KDS the target galaxies at $3 < z < 3.8$ were observed in both the K-band, into which the [O~{\sc III}]$\lambda$5007 and $H\beta$ lines are redshifted, and the H-band, into which the [O~{\sc II}]$\lambda$3727,$\lambda$3729 doublet is redshifted, to facilitate both dynamical and chemical abundance measurements.
The GOODS-S pointings were observed in the H and K-bands separately, however due to loss of observing time the SSA22 galaxies were observed with the KMOS HK filter, which has the disadvantage of effectively halving the spectral resolution, but allows for coverage of the H-band and K-band regions simultaneously. \\

We prepare each pointing using the KARMA tool \citep{Wegner2008}, taking care to allocate at least one IFU to observations of a `control' star closeby on the sky to allow for precise monitoring of the evolution of seeing conditions and the shift of the telescope away from the prescribed dither pattern (see \cref{subsubsec:datareduction}).
For the four pointings described above and summarised in table \ref{tab:pointings}, we adopt the standard object-sky-object (OSO) nod-to-sky observation pattern, with 300s exposures and alternating $0.2^{\prime\prime}$/$0.1^{\prime\prime}$ dither pattern for increased spatial sampling around each of the target galaxies.
This will later allow for datacube reconstruction with 0.1$^{\prime\prime}$ size spaxels as described in \cref{subsubsec:datareduction}. \\

The observations were carried out during ESO observing periods P92-P96 using Guaranteed Time Observations (Programme IDs: 092.A-0399(A), 093.A-0122(A,B), 094.A-0214(A,B),095.A-0680(A,B),096.A-0315(A,B,C)) with excellent seeing conditions.
In GOODS-S-P1 and GOODS-S-P2 the median K-band seeing was $\sim 0.5^{\prime\prime}$ and for the SSA22-CLUSTER and SSA22-FIELD pointings the K-band seeing ranged between $\sim 0.55-0.65^{\prime\prime}$.
We observed 19-21 targets in each field (see table \ref{tab:pointings}), with these numbers less than the available 24 arms for each pointing due to the combination of three broken pickoff arms during the P92/93 observing semesters and our requirement to observe at least one control star throughout an Observing Block (OB). \\

@@

\subsubsection{Data reduction}\label{subsubsec:datareduction}
The data reduction process primarly makes use of the Software Package for Astronomical Reduction with KMOS, (SPARK; \cite{Davies2013}), implemented using the ESO Recipe Execution Tool (ESOREX) \citep{Freudling2013}.
In addition to the SPARK recipes, custom python scripts were run at different stages of the pipeline and will be described throughout this section.

@@

Sky subtraction is enhanced using the SKYTWEAK option within SPARK \citep{Davies2007}, which counters the varying amplitude of OH lines between exposures by scaling `families' of OH lines independently to match the data.
Wavelength miscalibration between exposures due to spectral flexure of the instrument is also accounted for by applying spectral shifts to the OH families during the procedure, and in general the use of the SKYTWEAK option in the K-band greatly reduces the sky-line residuals. 
We monitor the evolution of the atmospheric PSF and the position of the control stars over the OBs, to allow us to reject raw frames where the averaged K-band seeing rises above $0.8^{\prime\prime}$ and to measure the spatial shifts required for the final stack more precisely.
The PSF is determined from fitting the K-band collapsed, stacked standard star cube in each pointing with an elliptical gaussian and the values are reported in table \ref{tab:pointings}.
The telescope tends to drift from its acquired position over the course of an OB and the difference between the dither pattern shifts and the measured position of the control stars provides the value by which each exposure must be shifted to create the stack. \\

@@

This analysis provides us with three crucial morphological parameters required to support the kinematic analysis of \cref{subsec:3d_modelling}, namely the axis ratio $b/a$, half-light radius $R_{1/2}$ and the morphological position angle $PA_{morph}$.
Example {\em HST} images and {\tt GALFIT} outputs are given in figure \ref{fig:main_body_kinematic_grids} for selected galaxies, with the full spatially resolved sample (see \cref{subsubsection:spaxel_fitting}) plotted throughout \cref{app:kinematics_plots}.

@@

b012141\_012208 & 03:32:23 & -27:51:57 & 3.471        & 24.12  & 10.16 & 59.0 & 0.36        & 72.0        & 9.0     & 1.57      \\
b15573          & 03:32:28 & -27:51:00 & 3.583        & 23.6   & 10.37 & 27.0 & 0.28        & 78.0        & 146.0   & 0.52      \\

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
%\author[O.J. Turner et al.]{
%O. J. Turner,$^{1,2,\thanks{E-mail: turner@roe.ac.uk (OJT)}}$
%R. K. Cochrane,$^{1}$
%M. Cirasuolo,$^{1,2}$
%C. M. Harrison,$^{2,3}$
%J. Dunlop,$^{1}$\newauthor
%R. J. McLure,$^{1}$
%A. M. Swinbank$^{4,3}$
%H. L. Johnson$^{3,4}$
%\\
%$^{1}$SUPA\thanks{Scottish Universities Physics Alliance}, Institute for Astronomy, University of Edinburgh, Royal Obervatory, Edinburgh EH9 3HJ\\
%$^{2}$European Southern Observatory, Karl-Schwarzschild-Str. 2, 85748 Garching b. M{\"u}nchen, Germany\\
%$^{3}$Centre for Extragalactic Astronomy, Durham University, South Road, Durham, DH1 3LE, U.K.\\
%$^{4}$Institute for Computational Cosmology, Durham University, South Road, Durham, DH1 3LE, U.K.
%}

@@

% Abstract of the paper
\begin{abstract}
We present dynamical measurements for 77 star-forming galaxies at $z\sim3.5$ in the mass range $9.5 < log(M_{\star}[M_{\odot}]) < 10.5$ from the KMOS (K-band Multi-Object Spectrograph) Deep Survey (KDS).
These measurements constrain the internal dynamics, the intrinsic velocity dispersions $\sigma_{int}$ and rotation velocities, $V_{C}$, of galaxies in the high-redshift Universe.
We have made careful comparisons to other IFS samples which have been selected consistently, span a wide redshift baseline and use ionised gas emission to trace galaxy dynamics. 
The mean dispersion value of the galaxies in our sample is $\sigma_{int} = 70.8^{+3.3}_{-3.1} kms^{-1}$, consistent with the increase of $\sigma_{int}$ with redshift reported up to $z\sim3$ in other samples.
However only $34 \pm 8\%$ of our galaxies are rotationally-dominated ($V_{C}/\sigma_{int}$ > 1), with the sample average $V_{C}/\sigma_{int}$ value much less than at lower redshift.
The rotation dominated KDS galaxies show no clear offset from the local rotation velocity-stellar mass (i.e. $V_{C}-M_{\star}$) relation, although a smaller fraction of the galaxies are on the relation due to the increase in the dispersion dominated fraction.
These observations are consistent with a simple equilibrium model picture, in which random motions are boosted in high-redshift galaxies by the increase in gas fractions, accretion efficiency, specific star-formation rate and stellar feedback.
\end{abstract}

@@

The galaxy population at all redshifts appears to be bimodal in many physical properties \citep[e.g. as described in][]{Dekel2006}, with a preference for the most massive galaxies to lie in the red sequence, characterised by red optical colours, low Star Formation Rates (SFRs) and spherical morphologies, and less massive galaxies in the blue sequence, characterised by blue colours, high star formation rates and disky morphologies. 
For these blue, star-forming galaxies (SFGs) there is a roughly linear correlation between SFR and stellar mass ($M_{\star}$) \citep[e.g.][]{Daddi2007,Noeske2007,Elbaz2007}, in the sense that galaxies which have already accumulated a larger stellar population tend to have higher SFRs.
This correlation, or `main sequence' underpins the `equilibrium model', in which the SFR of galaxies is regulated by the availability of gas, with outflows and accretion events sustaining the galaxy gas reservoirs in a rough equilibrium as the galaxy evolves \citep[e.g.][]{Dave2012,Lilly2013,Saintonge2013}.
The main sequence has been studied comprehensively, using multi-wavelength SFR tracers, between $0 < z < 3$ \citep[e.g.][]{Rodighiero2011,Karim2011,Whitaker2012,Behroozi2013b,Whitaker2014,Rodighiero2014,Speagle2014,Pannella2014,Sobral2014,Sparre2015,Lee2015,Schreiber2015,Renzini2015,Nelson2016}, showing evolution of the relation towards higher SFRs at fixed $M_{\star}$ with increasing redshift, reflecting the increase of the cosmic Star Formation Rate Density (SFRD) in this redshift range \citep{Madau_2014}.
At each redshift slice it has been suggested that galaxies on the main sequence evolve secularly as regulated by their gas reservoirs, and so selecting such populations offers the chance to explore the evolution of the physical properties of typical SFGs across cosmic time.
This assumes that high-redshift main sequence galaxies are the progenitors of their lower redshift counterparts, which may not be the case \citep[e.g.][]{Gladders2013,Kelson2014,Abramson2016b} and also assumes that we can learn about galaxy evolution (i.e. how individual galaxies develop in physical properties over time) by studying the mean properties of populations at different epochs.

The picture is complicated by the addition of major and minor galaxy mergers which can rapidly change the physical properties of galaxies \citep[e.g.][]{Toomre1977,Lotz2008,Conselice2011,Conselice2014} and the relative importance of in-situ, secular stellar mass growth vs. stellar mass aggregation via mergers is the subject of much work involving both observations and simulations \citep[e.g.][]{Robaina2009,Kaviraj2012,Stott2013,Lofthouse2016,Qu2016}. 
To account for the growing number density of quiescent galaxies from z$\sim 2.5$ to the present day \citep[e.g.][]{Bell2004,Faber2007,Brown2007,Ilbert2010,Brammer2011,Muzzin2013,Buitrago2013} there must also be some process which shuts off star-formation within main sequence galaxies (i.e. quenching) which, to explain observations, is a function of both mass and environment \citep{Peng2010,Darvish2016}.

Recent cosmological volume simulations are able to resolve the complex interplay of baryonic processes which are at work as galaxies evolve, and can track the development of individual galaxies from early stages, to maturity and through quenching \citep{Dubois2014a,Vogelsberger2014b,Schaye2015}.
Observations can aid the predictive power of such simulations by providing constraints on the evolving physical properties of galaxy populations.
The observed dynamical properties of galaxies contain information about the transfer of angular momentum between dark matter halos and the baryons and the dissipation of this angular momentum thereafter (through gravitational collpase, mergers and outflows \citealt{Fall1983,Romanowsky2012,Fall2013}), constituting an important set of quantities for simulators to attempt to match.
Developments in both Integral-Field Spectroscopy (IFS) instrumentation and data analysis tools over the last decade have led to the observation of two-dimensional velocity and velocity dispersion fields for large samples of galaxies of different morphological types, spanning a wide redshift range \citep[e.g.][]{Sarzi2005,Flores2006,Epinat2008,ForsterSchreiber2009,Cappellari2011,Gnerucci2011,Epinat2012,Croom2012,Swinbank2012,Swinbank2012a,Bundy2015,Wisnioski2015,Stott2016,Harrison2017,Swinbank2017} and when interpreted in tandem with high-resolution imaging data from the Hubble Space Telescope ({\em {\em HST}}) these data provide exquisite information about the range of physical processes which are driving galaxy evolution.
In particular, in recent years the multiplexing capabilities of KMOS have allowed for IFS kinematic observations for large galaxy samples to build up over much shorter timescales \citep{Sobral2013,Wisnioski2015,Stott2016,Mason2016,Harrison2017} providing an order of magnitude boost in statistical power over previous high-redshift campaigns.

SFGs generally appear to be increasing in random motions with increasing redshift between $0 < z < 3$, as traced by their observed velocity dispersions, $\sigma_{obs}$ \citep{Genzel2008,ForsterSchreiber2009,Law2009,Cresci2009,Gnerucci2011,Epinat2012,Kassin2012,Green2014,Wisnioski2015,Stott2016}.
This has been explained in terms of increased `activity' in galaxies during and before the global peak in cosmic SFRD \citep{Madau_2014}, in the form of higher specific star formation rates (sSFR) \citep{Wisnioski2015}, larger gas reservoirs \citep{Law2009,ForsterSchreiber2009,Wisnioski2015,Stott2016}, more efficient accretion \citep{Law2009}, increased stellar feedback from supernovae \citep{Kassin2012} and turbulent disk instabilities \citep{Law2009,Bournaud2007,Bournaud2016}, all of which combine to increase $\sigma_{obs}$ and complicate its interpretation.

There is also an increasing body of work measuring the relationship between the observed maximum rotation velocity of a galaxy, a tracer for the total dynamical mass, and its stellar mass, known as the stellar mass Tully-Fisher Relation (smTFR) \citep{Tully1977}, with surveys reporting disparate results for the evolution of this relation with redshift \citep[e.g.][]{Puech2008,Miller2011,Gnerucci2011,Swinbank2012,Simons2016,Tiley2016a,Harrison2017,Straatman2017}.
In both cases, systematic differences in measurement and modelling techniques at high-redshift, especially with regards to beam-smearing corrections, combine with our poor understanding of progenitors and descendants to blur the evolutionary picture which these surveys paint.

@@

\subsubsection{GOODS-S}\label{subsubsec:sample_selection_goods}
Two of the FIELD environment pointings (GOODS-S-P1;GOODS-S-P2; see Table \ref{tab:pointings}) are selected within GOODS-S region \citep{Guo2013}; accessible from the VLT and with excellent multi-wavelength coverage, including deep {\em HST WFC3 F160W} imaging with $ 0.06^{\prime\prime}$ pixel scale and $ \sim 0.2^{\prime\prime}$ PSF, which is well suited for constraining galaxy morphology \citep{Grogin2011,Koekemoer2011}.
We select targets from the various spectroscopic campaigns which have targetted GOODS-S, including measurements from VIMOS \citep{Balestra2010,Cassata2014}, FORS2 \citep{Vanzella2005,Vanzella2006,Vanzella2008} and both LRIS and FORS2 as outlined in \cite{Wuyts2009}.
These targets must be within the redshift range $3 < z < 3.8$, have high spectral quality (as quantified by the VIMOS redshift flag equal `3' or `4', and the FORS2 quality flag equal `A') and we carefully exclude those targets for which the [O~{\sc III}]$\lambda$5007 or H$\beta$ emission lines, observable in the K-band at these redshifts, would be shifted into a spectral region plagued by strong OH emission.
The galaxies which remain after imposing these criteria are distributed across the GOODS-S field, and we select two regions where $\sim 20$ targets can be allocated to the KMOS IFUs (noting that the IFUs can patrol a $7.2^{\prime}$ diameter patch of sky during a single pointing).
In GOODS-S-P1 and GOODS-S-P2 we observe 20 and 17 galaxies respectively.\footnote{We note that the number of observed galaxies in GOODS-S-P2 does not include two targets which were later found to have $z < 0.5$.}

\subsubsection{SSA22}\label{subsubsec:sample_selection_ssa}
A single CLUSTER environment pointing (SSA22-P1; see Table \ref{tab:pointings}) is selected from the SSA22 field, \citep{Steidel1998,Steidel2000,Steidel2003,Shapley2003}, which, as mentioned above, is an overdensity of LBG candidates at $z \sim 3.09$.
Thousands of spectroscopic redshifts have been confirmed for these LBGs with follow up observations using LRIS \citep{Shapley2003,Nestor2013}.
A combination of deep {\em B,V,R} band imaging with the Subaru Suprime-Cam \citep{Matsuda2004}, deep narrow band imaging at 3640$\AA$ \citep{Matsuda2004} and at 4977$\AA$ \citep{Nestor2011} and archival {\em HST ACS} and {\it WFC3} provides ancillary data in excellent support of integral field spectroscopy, albeit over a shorter wavelength baseline and with shallower exposures than in the GOODS-S field.
Fortunately at $z \sim 3.09$ the [O~{\sc III}]$\lambda$5007 line is shifted into a region of the K-band which is free from OH features and so for the CLUSTER environment pointing we fill the KMOS IFUs with galaxies located towards the centre of the SSA22 protocluster (SSA22-P1).

We also add a further FIELD environment pointing to the south of the main SSA22 spatial overdensity where the density of galaxies is typical (SSA22-P2).
In SSA22-P1 and SSA22-P2 we observe 19 and 21 galaxies respectively.
In summary we have chosen three FIELD environment pointings and a single CLUSTER environment pointing across GOODS-S and SSA22, comprising a total of 77 galaxies, as described in Table \ref{tab:pointings}. \footnote{Additional pointings in the COSMOS and UDS fields were originally scheduled as part of the GTO project, however 50$\%$ of the observing time was lost to bad weather during these visitor mode observations.}

\subsection{Observations and data reduction}\label{subsubsec:observations_and_dr}

Our data for the 77 KDS targets were observed using KMOS.
KMOS is a second generation IFS mounted at Nasmyth focal plane on UT1 at the VLT.
The instrument has 24 moveable pickoff arms, each with an integrated IFU, which patrol a region 7.2$^{\prime}$ in diameter on the sky, providing considerable flexibility when selecting sources for a single pointing.
The light from each set of 8 IFUs is dispersed by a single spectrograph and recorded on a 2k$\times$2k Hawaii-2RGHgCdTe near-IR detector, so that the instrument is comprised of three effectively independent modules.
Each IFU has 14$\times$14 spatial pixels which are 0.2$^{\prime\prime}$ in size, and the central wavelength of the K-band filter has spectral resolution of $R \sim 4200$.
\subsubsection{Observations}\label{subsubsec:Obs}
To achieve the science goals of the KDS the target galaxies at $3 < z < 3.8$ were observed in both the K-band, into which the [O~{\sc III}]$\lambda$5007 and $H\beta$ lines are redshifted, and the H-band, into which the [O~{\sc II}]$\lambda$3727,$\lambda$3729 doublet is redshifted, to facilitate both dynamical and chemical abundance measurements.
The GOODS-S pointings were observed in the H and K-bands separately, however due to loss of observing time the SSA22 galaxies were observed with the KMOS HK filter, which has the disadvantage of effectively halving the spectral resolution, but allows for coverage of the H-band and K-band regions simultaneously. \\

We prepare each pointing using the KARMA tool \citep{Wegner2008}, taking care to allocate at least one IFU to observations of a `control' star closeby on the sky to allow for precise monitoring of the evolution of seeing conditions and the shift of the telescope away from the prescribed dither pattern (see \cref{subsubsec:datareduction}).
For the four pointings described above and summarised in Table \ref{tab:pointings}, we adopt the standard object-sky-object (OSO) nod-to-sky observation pattern, with 300s exposures and alternating $0.2^{\prime\prime}$/$0.1^{\prime\prime}$ dither pattern for increased spatial sampling around each of the target galaxies.
This will later allow for datacube reconstruction with 0.1$^{\prime\prime}$ size spaxels as described in \cref{subsubsec:datareduction}. \\

The observations were carried out during ESO observing periods P92-P96 using Guaranteed Time Observations (Programme IDs: 092.A-0399(A), 093.A-0122(A,B), 094.A-0214(A,B),095.A-0680(A,B),096.A-0315(A,B,C)) with excellent seeing conditions.
In GOODS-S-P1 and GOODS-S-P2 the median K-band seeing was $\sim 0.5^{\prime\prime}$ and for the SSA22-CLUSTER and SSA22-FIELD pointings the K-band seeing ranged between $\sim 0.55-0.65^{\prime\prime}$.
We observed 19-21 targets in each field (see Table \ref{tab:pointings}), with these numbers less than the available 24 arms for each pointing due to the combination of three broken pickoff arms during the P92/93 observing semesters and our requirement to observe at least one control star throughout an Observing Block (OB). \\

@@

\subsubsection{Data reduction}\label{subsubsec:datareduction}

The data reduction process primarly makes use of the Software Package for Astronomical Reduction with KMOS, (SPARK; \cite{Davies2013}), implemented using the ESO Recipe Execution Tool (ESOREX) \citep{Freudling2013}.
In addition to the SPARK recipes, custom python scripts were run at different stages of the pipeline and will be described throughout this section.

@@

Sky subtraction is enhanced using the SKYTWEAK option within SPARK \citep{Davies2007}, which counters the varying amplitude of OH lines between exposures by scaling `families' of OH lines independently to match the data.
Wavelength miscalibration between exposures due to spectral flexure of the instrument is also accounted for by applying spectral shifts to the OH families during the procedure, and in general the use of the SKYTWEAK option in the K-band greatly reduces the sky-line residuals. 
We monitor the evolution of the atmospheric PSF and the position of the control stars over the OBs, to allow us to reject raw frames where the averaged K-band seeing rises above $0.8^{\prime\prime}$ and to measure the spatial shifts required for the final stack more precisely.
The PSF is determined from fitting the K-band collapsed, stacked standard star cube in each pointing with an elliptical gaussian and the values are reported in Table \ref{tab:pointings}.
The telescope tends to drift from its acquired position over the course of an OB and the difference between the dither pattern shifts and the measured position of the control stars provides the value by which each exposure must be shifted to create the stack. \\

@@

This analysis provides us with three crucial morphological parameters required to support the kinematic analysis of \cref{subsec:3d_modelling}, namely the axis ratio $b/a$, half-light radius $R_{1/2}$ and the morphological position angle $PA_{morph}$.
Example {\em HST} images and {\tt GALFIT} outputs are given in Figure \ref{fig:main_body_kinematic_grids} for selected galaxies, with the full spatially resolved sample (see \cref{subsubsection:spaxel_fitting}) plotted throughout \cref{app:kinematics_plots}.

@@

b012141\_012208 & 03:32:23 & -27:51:57 & 3.471        & 24.12  & 10.16 & 59 & 0.36        & 72        & 9     & 1.57      \\
b15573          & 03:32:28 & -27:51:00 & 3.583        & 23.6   & 10.37 & 27 & 0.28        & 78        & 146   & 0.52      \\
bs006516        & 03:32:15 & -27:50:46 & 3.215        & 23.94  & 9.85  & 14 & 0.50         & 61        & 146   & 1.91      \\
